\section{Ablation Study}

The ablation code path is already prepared in the repository, but the full set of runs is still in progress. This section therefore records the intended experimental questions and leaves explicit space for the corresponding quantitative results.

These ablations are not arbitrary extensions. Each one targets a specific component that is easy to isolate in the ConvMixer design: the patch-forming front-end, the expressiveness of channel mixing, and the dependence of the model on training-set size. Together, they are meant to answer whether ConvMixer works well because of the exact combination proposed in the paper or because a broader family of simple patch-and-convolution hybrids is already strong on CIFAR-10.

\subsection{Patch Embedding vs.\ Convolutional Stem}

This ablation replaces the original patch embedding stem with a more conventional convolutional stem that preserves the downstream tensor shape. The goal is to test whether the patch projection itself is an important ingredient of ConvMixer, or whether a standard convolutional front-end can recover most of the same behavior once the later mixing blocks are unchanged.

Our working hypothesis is that the original patch-style stem is likely part of what makes ConvMixer distinctive, but the performance gap may be modest if the replacement stem remains shape-compatible and parameter-comparable. Results will be added after runs complete.

\subsection{Pointwise Channel Mixing vs.\ Non-Linear Channel Mixing}

The base ConvMixer block uses a single $1 \times 1$ convolution for channel mixing. The non-linear channel mixing ablation replaces that stage with a deeper non-linear mixer while keeping the rest of the block structure fixed. This tests whether the original pointwise mixer is sufficient, or whether additional channel-wise capacity materially changes performance.

The expected outcome is that this variant may improve flexibility but does not necessarily improve accuracy enough to justify its added complexity on CIFAR-10. Results will be added after runs complete.

\subsection{Pointwise Channel Mixing vs.\ No Channel Mixing}

This ablation removes the channel mixing stage entirely by replacing it with an identity mapping. It is the cleanest way to test whether the pointwise mixing path is essential, because the spatial depthwise branch and the classifier head remain intact.

The expected outcome is a noticeable drop in accuracy, which would support the interpretation that ConvMixer requires both spatial and channel mixing rather than spatial processing alone. Results will be added after runs complete.

\subsection{Data-Size Robustness}

It is assumed that transformers require a bigger amount of data than convolutional networks, when trained from scratch. Not only this claim is supported by the existing research (such as \citep{vitcnnstudy}), there are papers dedicated to bridge this gap (such as \citep{bridginggap}).

In fact, it might be that the poor performance of DeiT-Tiny in the main experiment could be attributed to the fact that CIFAR-10 is already a relatively small dataset, compared to ImageNet that was used in the original paper, and other datasets commonly used to teach transformers image classification.

Our final ablation evaluates how convmixer trains when using  smaller amounts of training data, and how it compares to Resnet-18 and DeiT-Tiny in that respect. For this experiment, we take 50\%, 10\% and 1\% of training set of CIFAR-10, and train each of the models on the limited data for 50 epochs. At the same time, we use the whole test dataset.

The results are expected and similar to the main experiment outcome. Of course, the smaller the data, the worse the performance of a model. However, with 50\% of the training set, the fall in test accuracy is not so catastrophic, with models trailing behind by just about 6\% to 7\%. 

Something interesting can be observed when training at just 1\% of data: surprisingly, DeiT-Tiny starts with better accuracy after the first epoch (about 17\% compared to about 10\% for convolutional models) and its training looks more stable. Eventually, after about 20 epochs, the convolutional models do outperform DeiT-Tiny as in other runs.
